---
title: "Lab 5"
author: "Roshni Katrak-Adefowora"
date: "2/6/2021"
output: html_document
---

## Part 1: Time series wrangling & forecasting

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
library(here)
```

Read in energy data and convert to tsibble
```{r}
energy <- read_csv(here("data", "energy.csv"))

energy_ts <- energy %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)
```

Exploratory time series visualization
```{r}
ggplot(data=energy_ts, aes(x=date, y=res_total))+
  geom_line()+
labs(y="Residential energy consumption \n (Trillion BTU)")
```

Seasonplot
```{r}
energy_ts %>% 
  gg_season(y=res_total)+
  theme_minimal()+
  labs(x="month",
       y="residental energy consumption (trillion BTU)")
```

More ways to explore the data:
Subseries plot
```{r}
energy_ts %>% 
  gg_subseries(res_total)
```

Decomposition (by STL)

"STL is a versatile and robust method for decomposing time series. STL is an acronym for “Seasonal and Trend decomposition using Loess”, while Loess is a method for estimating nonlinear relationships."

Notice that it allows seasonality to vary over time (a major difference from classical decomposition, and important here since we do see changes in seasonality).
```{r}
#find STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))

#view components
#components(dcmp)

#visualize decomposed components
components(dcmp) %>% 
  autoplot()+
  theme_minimal()
```

Autocorrelation function (ACF)
```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot()
```

Forecasting by Holt-Winters exponential smoothing
```{r}
#create the model
energy_fit <- energy_ts %>% 
  model(ets = ETS(res_total ~ season("M"))) #M means multiplicative, A is additive, N is non-seasonal

#forecast using the model 10 years into the future
energy_forecast <- energy_fit %>% 
  forecast(h="10 years")

#plot just the forecasted values (with 80 and 96% CIs)
energy_forecast %>% 
  autoplot()

#plot original data + forecasted values
energy_forecast %>% 
  autoplot(energy_ts)
```

Assessing residuals
```{r}
#append predicted values (and residuals) to original energy data
energy_predicted <- broom::augment(energy_fit)

#plot actual energy values and predicted values atop them

ggplot(data=energy_predicted)+
  geom_line(aes(x=date, y=res_total))+
  geom_line(aes(x=date, y=.fitted), color="red")
```

Residuals should be uncorrelated, centered at zero, and ideally normally distributed
```{r}
#check for normal distribution and if centered at zero with histogram
ggplot(data=energy_predicted, aes(x=.resid))+
  geom_histogram()
```

